### 2017-06-16 manuscript/manuscript.Rmd
Edit  ## R Studio

---
title             : "Curating Research Assets in Behavioral Sciences: A Tutorial on the Git Version Control System"
shorttitle        : "Curating Research Assets in Behavioral Sciences"

author: 
  - name: Matti Vuorre
    affiliation: 1
    corresponding: yes
    address: 406 Schermerhorn Hall, 1190 Amsterdam Avenue MC 5501, New York, NY 10027
    email: mv2521@columbia.edu
  - name          : "James P. Curley"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Columbia University, New York, USA"

abstract: >
  Recent calls for improving the reproducibility of behavioral sciences have increased attention to the ways in which researchers curate, share and collaborate on their research assets. However, these discussions have largely failed to provide practical instructions on how to do so. In this tutorial paper, we explain why version control systems, such as the popular Git program, are especially suitable for these challenges to reproducibility. We then present a tutorial on how to use Git from the computer's command line, and with the popular graphical interface contained in the R Studio development environment. This tutorial is especially written for behavioral scientists with no previous experience with version control systems. Git is easy to learn, presents an elegant solution to challenges to reproducibility related to curating research assets, facilitates multi-site collaboration and productivity (by allowing multiple collaborators to work on the same source simultaneously) and can be implemented on common behavioral science workflows with little extra effort. Git may also offer a suitable solution to transparent data (and material) sharing through popular online services, such as GitHub and Open Science Framework.

keywords          : "reproducibility; version control; git; research methods; open science"
wordcount         : "X"
bibliography      : ["references.bib"]
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
lang              : "english"
class             : "doc, 12pt, draftfirst"  # man / jou / doc
output            : papaja::apa6_pdf
header-includes:
   - \usepackage{tcolorbox}
---

```{r include = FALSE}
library(knitr)  # Make sure that knit directory is project root
opts_chunk$set(
    fig.align = "center"
)
library(papaja)
library(tidyverse)
```

# Introduction

The lack of reproducibility is increasingly recognized as a problem across scientific disciplines, and calls for changing the scientific workflow to enhance reproducibility have been published in a wide range of research areas, including Biology [@markowetz_five_2015], Ecology [@ihle_striving_2017], Neuroscience [@eglen_toward_2017] and Psychology [@munafo_manifesto_2017]. However, although calls to focus on reproducibility are now commonplace on the pages of leading scientific journals [e.g. @baker_1500_2016], only a small minority of researchers are familiar with the tools and practices that enable implementing reproducibility in their workflows. Therefore, although there now is a broad consensus that efforts to improve reproducibility are important, materials instructing researchers in using them are lacking.

## Reproducibility

Consider for a while if you have ever struggled with the following questions [@ihle_striving_2017, p.2]: Have you ever found a mistake in your results without knowing what caused it? Forgot what analyses you have already done and how (and possibly, why)? Lost datasets or information about the cases or variables in a dataset? Struggled in redoing statistical or computational analyses when new data became available? Had difficulty understanding what data to use or how in a project that you inherited from another researcher? Answering any of these questions in the affirmative suggests that your work might benefit from improving reproducibility [@ihle_striving_2017].

But what exactly is reproducibility? Reproducibility can be defined as follows: "A research project is computationally reproducible if a second investigator (including you in the future) can recreate the final reported results of the project, including key quantitative findings, tables, and figures, given only a set of files and written instructions." [@kitzes_practice_2017]. In the context of experimental Psychology, for example, a "project" could be an experiment or a set of experiments investigating a hypothesis, "reported results" could be figures or statistical analyses in a conference presentation or a manuscript.

This definition makes an important distinction between *reproducibility* and *replicability*. Methods and results sections in traditional journal articles have focused on ensuring replicability by giving detailed instructions on how to repeat the experiment and collect a data set similar to the original one. However, little or no emphasis is placed on issues related to reproducibility, such as details of the data processing pipeline and statistical model exploration, although these issues are often more important to the reliability of the scientific results than are their end results (e.g. reported *p*-values; @leek_statistics:_2015). To put it more generally, replicability is a broad conceptual issue concerning the epistemology of scientific claims. Reproducibility, on the other hand, is a technical requirement for allowing others (including researchers themselves in the future) to assess the reliability of a specific set of reported empirical or computational results.

Further, although they are sometimes erroneously conlated [@open_science_collaboration_estimating_2015; @peng_reproducible_2011] reproducibility and replicability are independent of one another. Although the distinction between the two may be somewhat blurred in purely computational areas (e.g. simulation studies; @peng_reproducible_2011), it is clear that an experiment can be replicated even if the materials of the original study are not available for replicability. Likewise, the materials of an experiment might allow its results to be reproduced, yet the underlying scientific hypothesis might not be replicable. Therefore, it is important to keep these two concepts distinct; in the remainder of the manuscript we explicitly only discuss reproducibility.

Given this broad definition of reproducibility, it is easy to recognize that many projects fall between the extremes of completely replicable and not replicable at all. A completely reproducible project would provide consumers the entire raw data set, and detailed, unambiguous instructions to analyze it (usually as computer code). This degree of reproducibility is very rare in Psychology [@wicherts_poor_2006; @vanpaemel_are_2015], although data sharing has become more common in recent years in journals that award authors who do so with special badges [@kidwell_badges_2016]. However, even when authors (report to) share data, the shared data sets sometimes turn out to be incorrect, not usable, incomplete, or not available at all [@kidwell_badges_2016]. This suggests that reproducibility requires more than good intentions. What, then, are some specific challenges to reproducibility?

## Challenges to reproducibility

Evidence suggests that the level of reproducibility within Psychology is not very high. One survey of leading Psychology journals found that 34-58% of articles published between 1985 and 2013 had at least one inconsistently reported *p*-value [@nuijten_prevalence_2015]: Upon recalculation, the authors couldn't obtain the same *p*-values from the reported test statistics, and therefore these 34-58% of articles were, to some degree, non-reproducible (the inferential statistics could not be reproduced).

Why, do so many research products fall short of the basic scientific standard of reproducibility? We suggest one reason is the poor level at which researchers organize, curate, and collaborate on the research assets---e.g. data and code used to analyze it---which support the product's conclusions. In one study, researchers asked authors of 141 Psychology articles (with a total of 249 studies) to share their data for reanalysis. 73% of the authors refused to share their data, leading the study's authors to suggest that the considerable efforts involved in cleaning datasets at the final stages of publication may make data sharing unattractive to many [@wicherts_poor_2006; see also @vanpaemel_are_2015]. Vanpaemel et al. [-@vanpaemel_are_2015] highlight some reasons provided by authors for not sharing their data:

>"To our surprise, some authors are apparently willing to share, but have no easy access to their own data or have lost their data altogether, due to computer crashes or collaborators having left the university. Many authors cite a lack of time as a reason not to share, and note that sharing their data would take too much effort, which is probably due to poor documenta- tion and storage practices." (p. 2-3)

In short, scientists usually excel at *producing* research assets, but commonly fail at *curating* them (a topic commonly known as Digital Asset Management; @jacobsen_implementing_2012). This state of affairs is not surprising because researchers are trained to do research, but receive little or no training in formal methods or accepted gold standards for curating their research assets. For simple[^simple] projects, such as writing a solo-author manuscript, this may not be a problem. However, with increasing complexity of the type of data collected by behavioral scientists, and the increased number of collaborators, research sites, and technical skills required in modern scientific workflows, not knowing how to curate research assets can lead to wasted time or worse problems, as detailed above.

Fortunately for the empirical sciences, challenges related to organizing and curating materials ("digital assets") across time, space, and personnel have been solved to a high standard in computer science with Version Control Systems (VCS). In the remainder of this article, we introduce a popular VCS called Git, and illustrate its use in the scientific workflow with a hypothetical example project. In the tutorial below, we show how to use the Git VCS by using text commands from the computer's command line. After introducing the fundamentals of Git's VCS functions from the command line, we also show how to use Git easily with a Graphical User Interface (GUI) implemented in the popular R Studio Integrated Development Environment (IDE; @rstudio_team_rstudio:_2016).

# Version Control Systems

Version Control Systems are computer programs designed for tracking changes to computer code, and collaborating on the code with others. Although initially developed for writing code collaboratively[^vcs-collaborate], it is easy to recognize that VCS can also be adopted for scientific production and collaboration. For example, behavioral experiments are often created with a programming language, and can have multiple authors and versions. Keeping track of the versions of the program and changes to the code, and allowing many authors to contribute to it (without breaking the experimental program) are problems that VCSs were specifically designed for. VCS is also easily adopted to other aspects of the research cycle, such as curating data across computers and laboratories, or writing manuscripts with multiple authors, versions, and sources of results.

The core concepts of version control are that contributors to a project create small checkpoints of the changes they make to the source code---analogous to saving an intermediate version of a file on the computer's hard disk---and then submit those changes to the VCS. The VCS maintains a history of changes to the code between these little checkpoints, and therefore allows coming back to any earlier version by browsing the history. See Figure \@ref(fig:git-diagram) for a diagram of the typical VCS workflow.

Some popular Version Control Systems are SVN, Mercurial, and Git. In this tutorial, we focus on Git: Git is already incresing in popularity in the scientific community, and is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.

<!-- Examples of problems in psychology solved by Git. -->
\begin{tcolorbox}[colframe=green!50!black, colback=green!5, fonttitle=\small\bfseries, fontupper=\footnotesize, title=Box 1. How Git facilitates the scientific workflow.]
    \begin{itemize}
    \item How to try different ways of visualizing and modeling data while keeping track of the different versions---and the differences between them?
        \begin{itemize}
        \item Git saves each version of the analysis, allowing testing new features without losing previous versions or proliferating files in the project's directories. All the past versions can be directly compared with each other.
        \end{itemize}
    \end{itemize}
    \begin{itemize}
    \item How to work on the same code or manuscript files simultaneously with collaborators?
        \begin{itemize}
        \item Git was designed for distributed collaborative work: Collaborators work on their local copies and "push" and "pull" material to and from each other or a central "repository". Git keeps precise track of who has done what, when, and (possibly) why.
        \end{itemize}
    \end{itemize}
    \begin{itemize}
    \item How to share my work in an organized manner with others?
        \begin{itemize}
        \item Git enforces a common organization scheme among collaborators, making it easier to keep everyone "on the same page" with what goes where, and how to contribute to specific parts of the project. Sharing the project with others is built into Git, and can be facilitated with online services such as GitHub.
        \end{itemize}
    \end{itemize}
\end{tcolorbox}

## The Git Version Control System

Git[^git-name], unlike the operating system's (OS) default file viewer (Mac's Finder, Windows' File Explorer), is not a point-and-click program for navigating files and folders on a computer. Instead, Git adds functionality to the existing file system, by making available a specific set of commands---either executed from the command line, or through a graphical user interface (GUI)---which allow keeping track of files and their history, and distributing the files across multiple computers and users. Because Git is a standalone program not usually included in standard OS installs, users must first download and install the Git software on their computers. Git is free and open source, works on Windows, Mac, and Linux OSs (among others), and can be freely downloaded at <https://git-scm.com/>.

## Installing Git

Even if you already have Git installed (some computers do) it is good practice to install the latest version[^git-version], which can be downloaded as a standalone program from <https://git-scm.com/download>. For Mac users, the easiest way to install or update Git to the latest version is to download the installer from <http://git-scm.com/download/mac>, and install it like any other program. Windows users can download the Git software installer from <http://git-scm.com/download/win>. Linux (and other) users can download Git and find further install instructions at the Git website (<https://git-scm.com/book/en/v2/Getting-Started-Installing-Git>).

## Git setup

Once you have installed Git, you can be use it from the command line or through a GUI--there are various GUIs for using Git, and below we will explain how to use the R Studio Git GUI. First, it is important to learn a few basic Git commands from the OSs command line. The command line is a text-based interface for interacting with your computer, and its functionality greatly extends that of the standard way of interacting with the computer by clicking and pointing with a mouse.

To access the command line interface, you need to use a command line "shell" application. Mac users can open the built-in app *Terminal*, and Windows users can download and use one of many alternatives, such as Microsoft PowerShell (<https://msdn.microsoft.com/powershell>). After opening the shell app, you can type in commands to access the interface. For example typing `ls` (hit Return to execute) lists the files in the current folder. Some users may find using a text-based command line interface unfamiliar, but to get started with Git, there are only two required configuration commands which you need to run once, and the basic functionality requires using only four commands[^cli-links]. However, users can also read through the first part of the tutorial without executing the commands and simply focus on the workflow; we show how to use Git with a GUI in the second part of the tutorial.

The first step in using Git is making Git aware of who is using the computer. You need to set the user name and email address by entering a few basic commands in the command line. First, to show the current user information, type `git config --global user.name`, and hit return. This should not return anything, unless a previous user of the computer has set the global Git user name.

Each Git command starts with the word `git`, then a command (such as `config`), and then arguments to the command, such as `--global` (for global configuration) followed by the `user.name` variable name to show the global user name. For detailed instructions on how to use Git commands, you can type `git --help`. For help on how to use the `git config` command, type `git config --help`.

To ensure that Git knows who you are, type `git config --global user.name "User Name"` (where User Name is your name) in the command line, and hit return. This command maps `"User Name"` to Git's global `user.name` variable. If you now re-run the first command (`git config --global user.name`), the command line will return the name you entered as `"User Name"`. The user name can be anything you'd like, but it is probably a good idea to use your real name so that potential collaborators know who you are. The second piece of information is your email address, which is entered by `git config --global user.email "email@address.com"` (where `"email@address.com"` is your email address). You can verify that the correct email address was saved by typing `git config --global user.email` in the Terminal, and hit return.

Once this information is entered, Git will know who you are, and is able to track who is doing what and when within a project, which is especially helpful when you are collaborating with other people, or when you are working on multiple computers.

# Using Git

The first operating principle of Git is that your work is organized into independent projects, which Git calls *repositories*[^git-link-submodules]. A repository is a folder on your computer which is version controlled by Git (you can tell if a folder is "monitored" by git by checking if the folder, or any of its parent folders, contains a hidden `.git` directory). Everything that happens inside a repository is tracked by Git, but you have full control of what is committed to Git's history and when. Because you have full control of what and when is committed to history, there is a small set of operations you need to know[^git-cheat-sheet].

Briefly, when you work in a Git repository (make changes to files within it), Git monitors the state of each file, and when they change Git knows that they differ from the previously logged state. If you are happy with the current changes, you **add** the changed files to Git's "staging area". If you then are certain that the changes in the staging area are desirable, you **commit** the changes. These two operations are the backbone of using Git to store the state of the project whenever meaningful changes are made. Importantly, each commit in the repository's history contains information to recover the full state of the project at that point in time. Users can always go back to an earlier version by **checking out** a previous state from Git's history. This core of the Git workflow is illustrate in Figure \@ref(fig:git-diagram).

```{r git-diagram, fig.cap = "A diagram illustrating the typical Git workflow. Verbs indicated in bold text are Git operations and explained in detail in the main text."}
knitr::include_graphics("../data/git-diagram.png")
```

To understand the Git workflow in practice, we now turn to a practical example using a hypothetical project. Git can be added to a project at any stage of the project's life cycle, but to most clearly show its use, we begin with an empty project with nothing in it.

## Organizing files and folders

Implementing reproducibility into the scientific workflow is less time-consuming and effortful if it is planned from the onset of the project, rather than added to the project after all the work has been completed. It is therefore important to organize the project keeping a few key goals in mind (here, we follow guidelines such as the Project TIER recommendations[^tier-link]): The files and folders should have easy to understand names (avoid idiosynchratic naming schemes), and the names should indicate the purposes of the files and folders.

The first step is to create a home folder for the project. This folder should have an immediately recognizable name, and should be placed somewhere on your computer where you can find it. We call the example project `git-example`. All the materials related to this project will be placed in subfolders of the home directory, but for now `git-example` is just an empty folder waiting to be filled with data, code, and documents. To make the example concrete, you can follow the tutorial by typing out the following commands on your own computer.

## Initializing a Git repository

Next, you need to create a new folder for the Git repository. First, choose an appropriate folder on your computer (such as `User/Documents/`) where you'd like to create the project. You can either use the system's file navigator (Finder / File Explorer) to create this folder, or the following commands in the command line: Navigate to the desired folder by using `cd Documents` to move into the `Documents` folder (assuming it exists in the folder where you currently are), and `cd ..` to move out of the folder (to its containing folder.) Once you are in the folder where you want to create the project, type `mkdir git-example` to make the `git-example` directory, then `cd git-example` to move into it. Once you are in the project's home folder (you can verify where you are by typing `pwd`), you want to turn the folder into a project by initializing Git:

```bash
> git init
```

This command initializes the folder as a Git repository, and the only change so far has been the addition of a hidden `.git` folder inside `git-example` (and a `.gitattributes` file. Users can ignore these hidden files and folders, however they can be shown by typing `ls -la`.) Now that the folder is initialized as a Git repository, Git monitors any changes within it, and allows you to add and commit these changes (Figure \@ref(fig:git-diagram).)

## Adding a file to Git

Every project (repository) should contain a brief note explaining what the project is about and who to contact. This note is usually called a readme, and therefore our first contribution to this project will be a README file (this file is so important that it has become standard practice to write it in capital letters). The README file should be a plain text file (i.e. not created with Microsoft Word) that can be read with a simple text editor. You can create this file with any text editor (Macs have TextEdit app installed by default, which allows creating plain text files). We created the file with some text in it, and it can now be seen in the `git-example` folder, either by using the system file viewer or `ls` in the command line. Because we added this file to a Git repository, Git is also aware of it. To see what files have changed since the last status change in the repository (there clearly has been only this one), you can ask for Git's **status**:

```bash
> git status

On branch master
Initial commit
Untracked files:
  (use "git add <file>..." to include in what will be committed)
	README
nothing added to commit but untracked files present (use "git add" to track)
```

The relevant output returned from executing this command is the "Untracked files:" part. There, Git tells that there is an untracked file (README) in the repository. To start tracking changes in this file, we **add** it to Git by using the command `git add` followed by either a `.` (for adding all untracked files) or `README` for only tracking the `README` file.

```bash
> git add README
```

We've now added this file to the staging area (Figure \@ref(fig:git-diagram)), and if we are happy with changes to the file's status, we can **commit** the file to Git's history. Commits are essential Git operations: They signify meaningful changes to the repository, and can be later browsed and compared to one another. As such, it is helpful to attach a small message to each commit, describing why that commit was made. Here, we've created a README file, and our commit command would look as follows:

```bash
> git commit -m "Add README file."
```

The quoted text after the `-m` argument is the *commit message*. Entering this command to the command line returns a brief description of the commit, such as how many files changed, and how many characters inside those files were inserted and deleted.

## Keeping track of changes to a file with Git

The `git-example` project (or rather, Git) now keeps track of all and any changes to README. To illustrate, you can change the text in the README file with a text editor, save the file, and then ask for Git's status with `git status` on the command line:

```bash
> git status

On branch master
Changes not staged for commit:
  (use "git add <file>..." to update what will be committed)
  (use "git checkout -- <file>..." to discard changes in working directory)
	modified:   README
no changes added to commit (use "git add" and/or "git commit -a")
```

Git can tell that the README file has changed. It is often useful to know exactly *how* a file has changed, before committing it. To view **diff**erences to a file not yet committed, use `git diff <file>`. It shows changes within `<file>`, line by line, highlighting removals with red and added lines with green. Once you are happy with the changes, you can repeat the add and commit steps from above to permanently record the current state of the project to Git's log (below we use the `.` shortcut for adding all changes):

```bash
> git add .
> git commit -m "Populate README with project description."
```

## What does Git know?

The real importance of these somewhat abstract steps becomes apparent when we consider the Git **log**. To reveal the commit log of your repository, call:

```bash
> git log
```

The main outputs of this command show that each commit is identified with a unique hash code (long alphanumeric string), which we can use to call for further information (see below); an author; a date and time; and the short commit message. Executing `git log` on our example project at this stage returns this:

```bash
commit 60cbe5c9b4a78e500314f791080381030577a035
Author: Matti Vuorre <mv2521@columbia.edu>
Date:   Tue Jun 13 17:20:27 2017 -0400
    Populate README with project description.

commit 16c475023ecbc99446164187eeaaab10647ac550
Author: Matti Vuorre <mv2521@columbia.edu>
Date:   Tue Jun 13 17:14:14 2017 -0400
    Add README file.
```

To see what exactly changed in the last commit (the log has latest commits at the top), you can call `git show` with the commit's hash (only relevant parts of output shown below):

```bash
> git show 60cbe5c9b4a78e500314f791080381030577a035
commit 60cbe5c9b4a78e500314f791080381030577a035
Author: Matti Vuorre <mv2521@columbia.edu>
Date:   Tue Jun 13 17:20:27 2017 -0400
    Populate README with project description.
diff --git a/README b/README
--- a/README
+++ b/README
@@ -0,0 +1,8 @@
+# Example Git Project
+This example project illustrates the use of Git.
+authors:
+Matti Vuorre <mv2521@columbia.edu>
+James Curley
+2017
```

This output is a detailed log of all changes in that commit. From top, it lists the author, the message, and then the commit's "**diff**" (i.e. what differs in the new version vs the old version of the file.) The current diff shows that 1 file received eight additional lines of text (the part wrapped in @ symbols), and then the additions themselves (the lines prepended with +s).

Although you have now seen the fundamentals of using Git to track the states of (and therefore changes to) a repository, this contrived and overly simplistic example doesn't allow full appreciation of the benefits of using Git for version control. To better illustrate Git's functioning, we now fast-forward in the hypothetical example project to a stage in the project where more files and materials have been created.

## (Slightly More) Advanced Git

For example, after working for a while on the project, you could have added two files inside well-organized subfolders of the project. The project would like this when viewed with a file explorer:

```bash
git-example/
|-- admin/
    |-- ethics-info.pdf
|-- experiment-1/
    |-- analysis/
        |-- plan_n.R
|-- manuscript/
|-- README
```

Running `git status` now tells that there are two new files (possible empty folders are ignored). That is, since the previous commit, two files have been added to the project; one a pdf with some administrative information (`ethics-info.pdf`), the other one is an R [@r_core_team_r:_2017] script for a pre-planned power analysis (`plan_n.R`). You would probably like to track any changes to the power analysis script, but the ethics information file isn't really something that you need to keep track of---at least for this example. It would be laborious to constantly keep ignoring it when committing changes to Git (especially if we'd like to commit multiple files simultaneously with `git add .`) Git has an elegant solution to specifying which files to keep track of. Because by default all files are monitored, Git uses a special file for instructing which files are to be *ignored*.

### Make Git Ignore Files

To make Git ignore files, you simply add a plain text file called `.gitignore` to the home folder of the repository. You can use any text editor to create this file. Each row of this file should specify a file or a folder (or a regular expression) that Git should ignore. In the current example you could make Git ignore the `admin/` folder entirely, and any file with the .pdf extension inside `manuscript/`. The example .gitignore file would look like this (lines beginning with `#` are comments):

```bash
# Ignore everything inside the 'admin' folder
admin/

# Ignore all .pdf files in the 'manuscript' folder
manuscript/*.pdf

# Ignore a specific file in a specific folder
analysis/sandbox.R
```

Re-running `git status` now only shows the `plan_n.R` file (and the newly created .gitignore file, which is also under version control, naturally.)

Because there are now two untracked files that are not in .gitignore (`.gitignore` and `plan_n.R`), and you usually should aim to maintain a clean history for the project, you can create two separate commits: One for the Git ignore file, and one for the power analysis file.

```bash
git add .gitignore
git commit -m "Added .gitignore file"

git add .
git commit -m "Completed power analysis"
```

After this last commit, you can at any time come back to this commit with `git log` or `git show` and see what was inside the newly created power analysis file when it was first created. For instance, if new information suggests that you should change the assumed effect size in the power analysis, you can simply edit and save the file, then add and commit the changes to Git with a helpful message that logs this important event in Git's history.

This possibility of "rewinding history" is especially useful for files that might go under multiple revisions (manuscripts, analysis files), or if you are interested when and in what order the files were created---and who created or changed them. One might consider committing a power analysis as a small personal pre-registration of a part of the research plan.

### Try a new feature

We often find that making some changes to a project didn't have the desired effect: The manuscript ended weaker or the analysis didn't work anymore. Git allows great flexibility in trying new features, then undoing the changes[^git-branches]. Starting with an empty staging area, you could start modifying a file (e.g. `plan_n.R`), and after a while find out that the changes were not for good. At this point it is common to press "Undo" in the text editor, but if the file has been saved multiple times or multiple files have been changed, it is difficult to get to the starting point by simply using the "Undo" button. Instead, with Git you can **checkout** the file's previous version from history. To undo all changes to `plan_n.R` (since the last commit), run

```bash
git checkout experiment-1/analysis/plan_n.R
```

Notice that you have to write the full path of the file (relative to the project's root) so Git knows precisely which file you want to checkout from history. With these example operations, we have discussed the main Git operations as outlined in Figure \@ref(fig:git-diagram): Make changes to files, **add** to staging area, **commit** to history; **checkout** from history to undo changes.

# Collaborating

The true advantages of using Git become apparent when we consider projects with more than one contributor. For example, consider a project where data is collected at multiple sites, and these files are then saved onto a centralized server, or shared through a service that automatically merges files from multiple sources (such as the popular Dropbox service). If two or more sites accidentally save a data file with the same name (e.g. `data-001.csv`), and these changes are automatically merged, the later file will simply overwrite the earlier file. Disaster!

Alternatively, consider a data analysis where two or more people work simultaneously on some complicated analysis. If user A and B are making changes to the same file and user B saves the file, user A's version of the file will be overwritten. Disaster!

Git and other VCSs, on the other hand, were specifically designed to allow (and facilitate) multi-site collaboration on arbitrarily complex projects: Microsoft Windows is developed on a Git platform. Git is especially helpful in scientific collaboration, a topic to which we turn next.

There are many ways in which a team could collaborate on a Git project (e.g. <https://www.atlassian.com/git/tutorials/comparing-workflows>); here we focus on a common one, called the "Centralized Workflow" where one virtual copy of the project is considered the central node, and all contributors interact with this central node, instead of directly with each other.

## Overview of the Centralized Git Workflow

In this workflow, a central project hosted on a research team's server, or an online service like GitHub (<https://github.com/>), is used to send and receive information from and to local projects. Because the word "project" can now refer to virtual copies of projects, they are instead commonly referred to as **repositories**. First, one user creates this central repository, then other users **clone** a local copy of it (see Table \@ref(tab:git-commands)). Contributors, including the one who created the central repository, then work on their local projects as detailed above, and after committing **push** their changes to the central repository. To get changes that other users have pushed to the central repository, contributors **pull** changes from it.

```{r git-diagram-collaborate, fig.cap = "A diagram illustrating the typical collaborative Git workflow with a remote repository (e.g. GitHub). Verbs indicated in bold text are Git operations and explained in detail in the main text.", out.width="6in"}
knitr::include_graphics("../data/git-diagram-collaborate.png")
```

Setting up a centralized Git repository on a research team's private server is relatively straightforward, but because the details vary from team to team, here we illustrate the Centralized Workflow using GitHub.

## GitHub

GitHub is one of the 100 most popular website worldwide, and hosts over 60 million software projects with a total of over 20 million users (<https://github.com/about>). We chose GitHub for collaboration because it is already the de-facto standard in VCS collaboration among scientists, it offers free private (see below) repositories for students, and repositories from GitHub can easily be connected to projects hosted on the Open Science Framework (<https://osf.io>).

Anyone wanting to use GitHub must first create a free user account at <https://github.com>. After the user account has been set up, and the user has logged in, creating a new repository at GitHub is self evident (there is a big green button labeled "New repository"). You must create the GitHub account with the same email address as you used above when configuring your local Git (or re-configure your local Git to use your the email address that you used to register for GitHub).

### Create a new GitHub repository

Although we are creating a new repository here, the goal is to create an online "remote" repository for the running example within this tutorial. Because of this, do not check the "Initialize this repository with a README" (you already created one in the local repository), do not add a .gitignore or a license. For the current example, after clicking "New repository", we simply enter a name, which can be anything but for consistency we call the GitHub repository `git-example`---the same name as our local project folder. After entering the name, click "Create repository". The next step is to link the new remote to the local repository, and the required steps are detailed on screen when you do this at GitHub:

```bash
git remote add origin https://github.com/<username>/<reponame>.git
```

Where `<username>` and `<reponame>` are the user's GitHub username, and the GitHub repository name. The correct address is visible in the GitHub page that opens after creating the repository. Once the connection is set up, we can **push** local changes to the GitHub remote:

```bash
git push -u origin master
```

The `-u origin master` are only required for the first push, as they set up the connection. Running this command will send your local repository to the GitHub repository. For pushing changes collowing this initial push, simply type `git push` after adding and committing locally. You have now created the remote Central repository, and other users can start contributing to it.

### Contributing to a Central (GitHub) repository

Once another team member has set up Git on their own computer, and signed up for GitHub, they need to **clone** the remote GitHub repository onto their local computer. To clone a repository, the new user must first navigate to an appropriate location on the computer where they would like to save the project on their computer, for example their `User/Documents` folder. Once the appropriate location is found, cloning will create a new folder for the repository inside this folder.

```bash
git clone https://github.com/<username>/<reponame>.git
```

To find out the correct link to enter to `git clone`, you can navigate your web browser to the repository's GitHub address (e.g. <https://github.com/mvuorre/reproguide-curate> for this tutorial's repository) and click the big green "Clone or download" button; the complete address is in the text box.

New contributors can then work on their local copies as detailed in earlier parts of this tutorial; making changes, adding, committing, and pushing. After committing their changes, they can update the status of the Central Git repo by pushing their changes to it:

```bash
git push
```

### Obtaining other's changes from Central repo

Just as you must manually push your own local changes to the remote repository, you must also obtain others' changes by **pulling** them from the central repo. Pulling is indicated as the first step in the collaborative workflow in Figure \@ref(fig:git-diagram-collaborate), because it is important that you start working on the most up to date version of the project (e.g. you don't want to reinvent the wheel or make unnecessary conflicting changes). Before starting to work on your proposed changes, pull the remote changes with:

```bash
git pull
```

The way in which users and their local repositories interact with the central repository by pushing and pulling is the cornerstone of collaboration on GitHub, and thoughtful use of these commands allows for complex workflows without any important code (data, ideas in manuscript, analysis code) ever being overwritten. However, there is no automatic way for a computer to tell what changes to prioritize: If two or more users have worked on the same code and then attempt to push their changes, it is possible that they have made changes that conflict with each other. If this happens, and it will, there is no need to worry; you simply need to know how to resolve the conflict.

### Resolving a conflict

If user A has pushed changes to the central repository while user B was working on the same code, user B's later push can result in conflict. That is just a natural consequence of two individuals working simultaneously on the same idea, and then writing different code in the same location in the file. When this happens, user B needs to first pull the most up to date changes from the remote repository to her local code, and then inspect the code for problems (which will be apparent, Git will insert pointers to where the conflict is occurring.) Once user B's local code is updated with the pull, and the conflict manually resolved, user B can push changes from her local repository to the central repository.

How these potential conflicts appear depends on how users collaborate with one another, and a detailed explanation of all potential scenarios is outside the scope of this tutorial. For more information, see e.g. <https://www.atlassian.com/git/tutorials/comparing-workflows>. The GitHub customer service is also very responsive to users' help requests. Most importantly, even in the event of merge conflicts, all committed changes are saved in Git's history and can be retrieved.

## Private or public collaboration?

By default, all GitHub repositories are public: Anyone with an internet connection can use their web browser to inspect the contents of your repository, or even clone it to their computer. This may sound unfamiliar to researchers used to working more privately, and clearly necessitates planning and thought with respect to issues such as data privacy and sharing sensitive materials. However, for many projects---including the writing of this tutorial---we see very few downsides to working "in the open".

There are two alternatives to working in a public GitHub repository: One, which we won't cover here, is to not use GitHub but instead place the central repository on the research team's shared but private server. The second option is to make the repository private on GitHub (this can be done when the repository is first created or afterwards by clicking Settings on the repository's website). Private repositories, and their contents, are only accessible to invited team members, and are therefore ideal for small teams who would like to work without revealing their master plans to the public just yet. For example, you might initially choose to work in a private repository, and only make it public once you feel the material is mature enough for public consumption.

To make a GitHub repository private, navigate to the repository's website with a web browser, and click "Settings", then "Make this repository private". Once one user has set the central GitHub repository to private mode, anyone wishing to clone, push, pull, or view the repo must provide their GitHub username and password. Only if they match an invited team members username and password can the user access the repository.

# Using Git from IDE

Above, we have detailed how to use Git from the computer's command line, but not all users are comfortable---although there is no reason not to be!---with this mode of interacting with their computers. We chose to introduce Git from the command line because eventually users might want to use it for more complicated operations (and it is required for setting the user's information). However, there are various graphical user interfaces (GUIs) available for Git which allow point-and-click interacting with Git.

## R Studio

In this section, we introduce how to use the popular R Studio IDE [@rstudio_team_rstudio:_2016] for the R programming language [@r_core_team_r:_2017] for interacting with Git. An IDE is an interface that bundles together many necessary features of software development---some users may be familiar with R Studio for conducting statistical analyses. Version control is an essential feature of software development, and R Studio provides a good GUI within its IDE for controlling Git.

For this example, we will begin a completely new project using only R Studio, and assume that readers are familiar with the concepts from the above tutorial. R Studio is a free and open source IDE, works on Windows, Mac, and Linux operating systems, and can be downloaded from the project's website at <https://www.rstudio.com/products/rstudio/download/>. Once downloaded and installed, it can be used for accessing the R programming language, and for many operations involved in curating research assets, including a GUI for Git.

Managing projects (and Git repositories) with R Studio is centered on the idea of R Projects. To start a new project with R Studio, open the R Studio application, and click File -> New Project. This brings up a dialog asking whether to create a project in a new directory, existing directory, or checkout an existing project from a version control repository. Here, we create a new project in a new directory, and choose "New Project" in the following screen. Then we'll give a name to the project's home folder and choose where to save it on the computer. Importantly, we'll also check the "Create a git repository" box, which will automatically set up a new repository for the project (provided you have set Git up as detailed above). Clicking "Create project" creates the folder in the specified location, and two files inside the project's main folder.

```{r rstudio-project-1, fig.cap = "Creating a project in R Studio."}
knitr::include_graphics("images/rstudio-project-1.png")
```

One of these files is `.gitignore` which we covered above. The other file is and .Rproj file, which indicates that the folder is the home folder for an R (Studio) project. Users don't interact with this file, but when opened, it is simply a plain text file containing the project's settings (these can be modified through Tools -> Project Options in R Studio).

## Using Git with R Studio

Once the R Project has been created, R Studio has a "Git" tab in the top-right panel of the GUI. At first, this tab shows the two new files in the repository, and some buttons with familiar looking names ("Commit", "Push", etc.)

```{r rstudio-git-1, fig.cap = "R Studio's Git tab for a newly created project."}
knitr::include_graphics("images/rstudio-git-1.png")
```

### Add and Commit changes

To mark this milestone of creating a project, let's commit these changes to Git. The workflow is exactly the same as when using Git from the terminal, but we now have some visually appealing helpers. To begin committing these changes, click on "Commit" in the Git tab. This brings up another window where we can select files to add into the staging area (the `.gitignore` file in Figure \@ref(fig:rstudio-git-2) is staged by checking the box). Once all the desired files (here we chose both) are staged, we write a short commit message as detailed above, and click the "Commit" button.

```{r rstudio-git-2, fig.cap = "Creating a Git commit with R Studio."}
knitr::include_graphics("images/rstudio-git-2.png")
```

### Add a remote GitHub repository

When creating the R Project, there was an option for creating the project from an existing Git repository. Because we didn't do this, we must now manually instruct Git that this repo should have a remote in GitHub. This is done exactly as above, and we must use command line functions. After you've created a new repository on GitHub, navigate to the project's folder on your computer with the command line navigator of your choice, and execute the following commands:

```bash
git remote add origin https://github.com/<username>/<reponame>.git
git push -u origin master
```

After executing these commands, you can use all Git and GitHub features from R Studio without having to use the command line.

### Further examples

- Born open data [@rouder_what_2015]

# Git Summary

## Overview of Git Commands

```{r git-commands, results = 'asis', warning = F}
commands <- read_csv("data/git-commands.csv")
commands %>% 
    apa_table(
        caption = "Main Git commands.",
        note = "Source: https://help.github.com/articles/github-glossary/ and https://services.github.com/on-demand/downloads/github-git-cheat-sheet.pdf",
        landscape = FALSE)
```

<!-- Box 2. Further resources for learning Git -->
\begin{tcolorbox}[colframe=green!50!black, colback=green!5, fonttitle=\small\bfseries, fontupper=\footnotesize, title=Box 2. Further resources for learning Git]
\begin{itemize}
    \item \href{https://www.git-tower.com/blog/workflow-of-version-control}{Basic VCS workflow}, an infographic explaining how VCS works (\url{https://www.git-tower.com/blog/workflow-of-version-control}).
    \item \href{https://services.github.com/resources/cheatsheets/}{GitHub's Git cheat sheet} is available in multiple languages and contains the most used Git commands (\url{https://services.github.com/resources/cheatsheets/}).
    \item \href{https://try.github.io}{TryGit}, an interactive website for learning the basics of Git (\url{https://try.github.io}).
    \item \href{http://r-pkgs.had.co.nz/git.html}{Git + GitHub} in an R programming context (\url{http://r-pkgs.had.co.nz/git.html}).
    \item \href{https://git-scm.com/book/en/v2}{Pro Git book}, a complete manual of Git (\url{https://git-scm.com/book/en/v2}).
\end{itemize}
\end{tcolorbox}
\newpage

[^git-version]: As of the writing of this article, the current version of Git is 2.13.1.
[^git-name]: The creator of Git, Linus Torvalds, named Git after himself as "the stupid content tracker" (@mcmillan_after_2005; <https://git-scm.com/docs/git.html>)
[^simple]: By "simple", we only mean that the technical aspects of the project, related to curating materials related to it, are simple. We of course do not suggest that there might be anything theoretically or scientifically simple in such projects.
[^vcs-collaborate]: The software we present below is used used by major software developers such as Microsoft, Google, and Facebook on code bases with hundreds of contributors.
[^cli-links]: If using a text-based command line seems challenging, Codecademy (<https://www.codecademy.com/learn/learn-the-command-line>) has a free interactive online tutorial, and MIT offers a free online game to teach using the command line (<http://web.mit.edu/mprat/Public/web/Terminus/Web/main.html>).
[^git-link-submodules]: For advanced users, Git *submodules* allow linking projects to each other, or organizing more complex projects into projects and their sub-projects (<https://git-scm.com/book/en/v2/Git-Tools-Submodules>).
[^git-cheat-sheet]: It helps to have a Git command cheat sheet (<https://services.github.com/resources/cheatsheets/>) printed and taped on your wall, but it contains many more commands than are needed for the basic use of Git.
[^tier-link]: <http://www.projecttier.org/tier-protocol/specifications/>
[^git-branches]: A particularly powerful approach for trying new features is **branching**: The project can be duplicated to a new branch and modified, then merged back to the main branch after work on the new feature is complete---or the new branch can be discarded if the work ended unsatisfactory. Branches are outside the scope of this tutorial, for more information see the Git website (<https://git-scm.com/book/en/v2/Git-Branching-Branches-in-a-Nutshell>).

# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}